---
title: Infrastructure
author: Ada Zhou
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)

# Parameters

ROBINSON <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

file_data <- here::here("c01-own/data/infrastructure.rds")
map_file_path <- 
    here::here(
        "../../archive/ne_110m_admin_0_countries/ne_110m_admin_0_countries.shp"
    )
world_regions <- 
  c(
    "8S",
    "ZG",
    "Z4",
    "ZQ",
    "Z7",
    "ZJ",
    "XU"
  )

group_codes <- 
  c(
    "1A",
    "S3",
    "B8",
    "V2",
    "Z4",
    "4E",
    "T4",
    "XC",
    "Z7",
    "7E",
    "T7",
    "EU",
    "F1",
    "XE",
    "XD",
    "XF",
    "ZT",
    "XH",
    "XI",
    "XG",
    "V3",
    "ZJ",
    "XJ",
    "T2",
    "XL",
    "XO",
    "XM",
    "XN",
    "ZQ",
    "XQ",
    "T3",
    "XP", #MIDDLE INCOME
    "XU",
    "XY",
    "OE",
    "S4",
    "S2",
    "V4",
    "V1",
    "S1",
    "8S",
    "T5",
    "ZG",
    "ZF",
    "T6",
    "XT", #UPPER MIDDLE INCOME
    "1W"
  )

#===============================================================================

df <- read_rds(file_data)

countries <- read_sf(map_file_path)
```

### Checking Data

```{r}
df %>% 
  summary()

df %>% 
  count(
    electricity_urban_na = is.na(electricity_urban),
    electricity_rural_na = is.na(electricity_rural),
    water_urban_na = is.na(water_urban),
    water_rural_na = is.na(water_rural),
    sanitation_urban_na = is.na(sanitation_urban),
    sanitation_rural_na = is.na(sanitation_rural)
  ) 
```

### Tidying Data

```{r}
df <- 
  df %>% 
  drop_na(
    electricity_urban, 
    electricity_rural,
    water_urban,
    water_rural,
    sanitation_urban,
    sanitation_rural
  ) %>% 
  pivot_longer(
    cols = c(contains("urban"), contains("rural")),
    names_to = c("infrastructure", "u_r"),
    names_sep = "_",
    values_to = "pop_percent"
  )
```

### Mexico Data

```{r}
df %>% 
  filter(country == "Mexico") %>% 
  ggplot(aes(year, pop_percent, color = infrastructure)) +
  geom_point() +
  facet_wrap(vars(u_r))
```

### World Infrastructure Data

```{r}
#world
df %>% 
  filter(iso2c == "1W", year == "2017") %>% 
  ggplot(aes(x = infrastructure, y = pop_percent, fill = u_r)) +
  geom_col(position = "dodge")
```

```{r}

df %>% 
  filter(
    iso2c %in% world_regions
  ) %>% 
  ggplot(aes(year, pop_percent, color = country)) +
  geom_line() +
  facet_grid(rows = vars(infrastructure), cols = vars(u_r))

```


### Difference between urban and rural infrastructure in regions

```{r}

df %>%
  pivot_wider(
    names_from = c(u_r, infrastructure), 
    values_from = pop_percent
  ) %>% 
  mutate(
    e_diff = urban_electricity - rural_electricity,
    w_diff = urban_water - rural_water,
    s_diff = urban_sanitation - rural_sanitation
  ) %>% 
  filter(
    iso2c %in% world_regions
  ) %>% 
  drop_na(e_diff, w_diff, s_diff) %>%
  select(iso2c, country, year, contains("diff")) %>% 
  pivot_longer(
    cols = contains("diff"),
    names_to = "difference",
    values_to = "value"
  ) %>% 
  ggplot(aes(x = year, y = value, color = country)) +
  geom_line() +
  facet_grid(rows = vars(difference))
```

### Sanitation 

```{r, fig.height=10}

lowest_sanitation_rural <- 
  df %>% 
  filter(
    !(iso2c %in% group_codes), 
    infrastructure == "sanitation",
    u_r == "rural",
    year == max(year, na.rm = TRUE)
  ) %>% 
  top_n(-50, pop_percent) %>% 
  pull(iso2c)

df %>% 
  filter(
    iso2c %in% lowest_sanitation_rural,
    u_r == "rural",
    infrastructure == "sanitation"
  ) %>% 
  pivot_wider(
    names_from = year,
    names_prefix = "year_",
    values_from = "pop_percent"
  ) %>% 
  mutate(country = fct_reorder(country, year_2017)) %>% 
  drop_na(year_2007) %>%  
  ggplot(aes(y = country)) +
  geom_vline(aes(xintercept = median(year_2017)), color = "blue") +
  geom_vline(aes(xintercept = median(year_2007)), color = "red") +
  geom_segment(
    aes(
      x = year_2007, 
      xend = year_2017, 
      y = country, 
      yend = country
    )
  ) +
  geom_point(aes(x = year_2007, color = "2007")) +
  geom_point(aes(x = year_2017, color = "2017"))
  
```

```{r, fig.height=10}

lowest_sanitation_urban <- 
  df %>% 
  filter(
    !(iso2c %in% group_codes), 
    infrastructure == "sanitation",
    u_r == "urban",
    year == max(year, na.rm = TRUE)
  ) %>% 
  top_n(-50, pop_percent) %>% 
  pull(iso2c)

df %>% 
  filter(
    iso2c %in% lowest_sanitation_rural,
    u_r == "urban",
    infrastructure == "sanitation"
  ) %>% 
  pivot_wider(
    names_from = year,
    names_prefix = "year_",
    values_from = "pop_percent"
  ) %>% 
  mutate(country = fct_reorder(country, year_2017)) %>% 
  drop_na(year_2007) %>%  
  ggplot(aes(y = country)) +
  geom_vline(aes(xintercept = median(year_2017)), color = "blue") +
  geom_vline(aes(xintercept = median(year_2007)), color = "red") +
  geom_segment(
    aes(
      x = year_2007, 
      xend = year_2017, 
      y = country, 
      yend = country
    )
  ) +
  geom_point(aes(x = year_2007, color = "2007")) +
  geom_point(aes(x = year_2017, color = "2017"))
```

### Countries where percent population with access to infrastructure decreased  

```{r}

decreased <- 
  df %>% 
  unite("variable", infrastructure, u_r) %>% 
  filter(year %in% c("2007", "2017")) %>% 
  group_by(iso2c, country, variable) %>% 
  arrange(iso2c, country, variable, year) %>% 
  summarize(change = last(pop_percent, year) - first(pop_percent, year)) %>%
  ungroup() %>% 
  ungroup() %>% 
  filter(change < 0) 

graph_country <- function(iso2c, country) {
  df %>% 
    filter(
      iso2c == {{iso2c}}
    ) %>% 
    ggplot(aes(x = year, y = pop_percent, color = u_r)) +
    geom_line() +
    geom_point() +
    facet_grid(rows = vars(infrastructure)) +
    scale_x_continuous(breaks = scales::breaks_width(2)) +
    labs(
      title = str_glue("{country} data")
    )
}

decrease_plots <- 
  decreased %>% 
  filter(abs(change) > 1) %>% 
  distinct(iso2c, country) %>%
  pmap(graph_country)

decrease_plots %>% walk(print)

```


### Decreasing Infrastructure

```{r}

graph_variable_decrease <- function(variable) {
  df %>% 
    unite("variable", infrastructure, u_r) %>% 
    filter(
      variable == {{variable}},
      year > 2006,
      iso2c %in% 
        (decreased %>% 
           filter(variable == {{variable}}, abs(change) > 1) %>% 
           pull(iso2c)
        ),
      !(iso2c %in% group_codes)
    ) %>% 
    ggplot(aes(year, pop_percent, color = country)) +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = scales::breaks_width(1)) +
    labs(
      title = str_glue("Countries with decreasing {variable}")
    )
}

graph_variable_decrease("electricity_rural")
graph_variable_decrease("electricity_urban")
graph_variable_decrease("water_rural")
graph_variable_decrease("water_urban")
graph_variable_decrease("sanitation_rural")
graph_variable_decrease("sanitation_urban")

decreased %>%
  filter(!(iso2c %in% group_codes)) %>% 
  count(variable) %>% 
  ggplot(aes(fct_reorder(variable, n), n)) + 
  geom_col() +
  labs(
    title = "Number of countries with a decrease in population percent access to infrastructure"
  )

```

### Mean Decrease by infrastructure

```{r}

decreased %>%
  filter(!(iso2c %in% group_codes)) %>%
  group_by(variable) %>% 
  summarize(mean = mean(change)) %>% 
  ggplot(aes(fct_reorder(variable, mean), mean)) + 
  geom_col()

df %>% 
  unite("variable", infrastructure, u_r) %>% 
  filter(year %in% c("2007", "2017")) %>% 
  group_by(iso2c, country, variable) %>% 
  arrange(iso2c, country, variable, year) %>% 
  summarize(change = last(pop_percent, year) - first(pop_percent, year)) %>% 
  filter(!(iso2c %in% group_codes)) %>%
  group_by(variable) %>% 
  summarize(mean = mean(change)) %>% 
  ggplot(aes(fct_reorder(variable, mean), mean)) + 
  geom_col()

```


```{r, fig.height=10}
# India
 df %>% 
  filter(
    iso2c == "IN"
  ) %>% 
  ggplot(aes(x = year, y = pop_percent, color = infrastructure)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(u_r)) +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  theme(legend.position = "bottom")

```


```{r}

top_bottom <- function(infrastructure, u_r, n) {
  df %>%
  filter(
    !(iso2c %in% group_codes),
    infrastructure == {{infrastructure}},
    year == 2017,
    u_r == {{u_r}}
  ) %>% 
  top_n(n, pop_percent) %>% 
  arrange(desc(pop_percent))
}

top_bottom("electricity", "urban", -10)
top_bottom("sanitation", "urban", -10)
top_bottom("water", "urban", -10)
top_bottom("electricity", "rural", -10)
top_bottom("sanitation", "rural", -10)
top_bottom("water", "rural", -10)
 
```

```{r, fig.height=20, fig.width=10}

graphs_top_bottom <- function(infrastructure, u_r, n) {
  df %>% 
  filter(
    iso2c %in% top_bottom({{infrastructure}}, {{u_r}}, n)$iso2c,
    u_r == {{u_r}}
  ) %>% 
  ggplot(aes(x = year, y = pop_percent, color = infrastructure)) +
  geom_line() +
  facet_wrap(vars(country), ncol = 2) +
  theme(legend.position = "bottom")
}

graphs_top_bottom("electricity", "rural", -10)
graphs_top_bottom("sanitation", "rural", -10)
graphs_top_bottom("water", "rural", -10)
graphs_top_bottom("electricity", "urban", -10)
graphs_top_bottom("sanitation", "urban", -10)
graphs_top_bottom("water", "urban", -10)

```

### Maps

```{r}

country_df <- 
  df %>% 
  pivot_wider(
    names_from = c(infrastructure, u_r), 
    values_from = pop_percent
  ) %>% 
  left_join(countries, ., by = c("ISO_A2" = "iso2c"))



maps_infrastructure <- function(infrastructure, year) {
  country_df %>% 
    filter(year == {{year}}) %>% 
    ggplot() +
    geom_sf(aes(fill = {{infrastructure}})) +
    scale_fill_viridis_c(labels = scales::label_percent(scale = 1)) +
    guides(
      fill = 
        guide_colorbar(
          barheight = 0.5,
          barwidth = 15,
          title = NULL
        )
    ) +
    theme_void() +
    theme(legend.position = "bottom") 
}

maps_infrastructure(electricity_urban, 2017) + 
  labs(title = "Percent of Urban Population with Access to Electricity")
maps_infrastructure(electricity_rural, 2017) +
  labs(title = "Percent of Rural Population with Access to Electricity")
maps_infrastructure(sanitation_urban, 2017) +
  labs(title = "Percent of Urban Population with Sanitation")
maps_infrastructure(sanitation_rural, 2017) +
  labs(title = "Percent of Rural Population with Sanitation")
maps_infrastructure(water_urban, 2017) +
  labs(title = "Percent of Urban Population with Access to Water")
maps_infrastructure(water_rural, 2017) +
  labs(title = "Percent of Rural Population with Access to Water")

```


### Correlation between variables

```{r}

df %>% 
  pivot_wider(
    names_from = u_r,
    values_from = pop_percent
  ) %>% 
  filter(year == "2017") %>% 
  drop_na(urban, rural) %>% 
  ggplot(aes(urban, rural)) +
  geom_abline(slope = 1, color = "light blue") +
  geom_point(alpha = 0.7) +
  facet_grid(cols = vars(infrastructure)) +
  coord_fixed()

region_country <-
  read_csv(file_region_country)

df %>%
  pivot_wider(
    names_from = u_r,
    values_from = population_percent
  ) %>%
  filter(
    !(iso2c %in% group_codes),
    year == "2017"
  ) %>%
  drop_na(urban, rural) %>%
  left_join(region_country, by = c("country" = "CountryName")) %>%
  ggplot(aes(urban, rural, color = GroupName)) +
  geom_abline(slope = 1, color = "light blue") +
  geom_point(alpha = 0.5, size = 1) +
  facet_grid(cols = vars(infrastructure)) +
  coord_fixed()

# region_country <- 
#   read_csv(file_region_country)
# 
# q2 %>% 
#   filter(
#     !(iso2c %in% group_codes)
#   ) %>%
#   drop_na(population_percent) %>% 
#   distinct(iso2c, country) %>% 
#   left_join(region_country, by = c("country" = "CountryName")) 

```

### Rural > Urban

```{r}

rural_greater <- 
  df %>% 
  pivot_wider(
    names_from = u_r,
    values_from = pop_percent
  ) %>% 
  filter(
    rural - urban > 1
  )

rural_greater %>% 
  distinct(iso2c, country, infrastructure) %>% 
  knitr::kable()

rural_greater %>% 
  distinct(iso2c, country, infrastructure) %>% 
  ggplot(aes(infrastructure)) +
  geom_bar()

rural_greater_plots <- 
  rural_greater %>% 
  distinct(iso2c, country) %>%
  pmap(graph_country)

rural_greater_plots %>% walk(print)


```



```{r, include=FALSE}
df %>% 
  unite("variable", infrastructure, u_r) %>% 
  filter(
    variable == "electricity_urban",
    year > 2006,
    iso2c %in% 
      (decreased %>% filter(variable == "electricity_urban") %>% pull(iso2c)) 
  ) %>% 
  ggplot(aes(year, pop_percent, color = country)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  coord_fixed(ylim = c(97, 100))

df %>% 
  unite("variable", infrastructure, u_r) %>% 
  filter(
    variable == "electricity_urban",
    year > 2006,
    iso2c %in% 
      (decreased %>% filter(variable == "electricity_urban") %>% pull(iso2c)),
    !(iso2c %in% group_codes)
  ) %>% 
  ggplot(aes(year, pop_percent, color = country)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  theme(
    legend.position = "bottom"
  )

```


```{r}
graph_country <- function(iso2c, infrastructure, u_r) {
  q2 %>% 
    filter(
      iso2c == {{iso2c}},
      infrastructure == {{infrastructure}},
      u_r == {{u_r}}
    ) %>% 
    ggplot(aes(x = year, y = population_percent)) +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = scales::breaks_width(2)) +
    labs(
      title = str_glue("{iso2c} data"),
      x = "Year",
      y = str_glue("Access to {infrastructure} (Percent of Population)"),
      captions = "Source: World Bank Data"
    )
}
```

__q4__ Graph the distribution of change for access to electricity, water and sanitation in urban and rural areas for every country. What information does the graphs tell you about the change in access to each infrastructure?

Hint: Use the vector `group_codes` to select only country-level data.

```{r, fig.height=10, fig.width=10}
# solution-begin

q4 <- 
  q2 %>% 
  filter(
    !(iso2c %in% group_codes),
    year %in% c("2007", "2017")
  ) %>%
  mutate(
    u_r = ordered(
      u_r,
      levels = c("urban", "rural"),
      labels = c("Urban", "Rural")
    ),
    infrastructure = 
      ordered(
        infrastructure, 
        levels = c("water", "electricity", "sanitation"),
        labels = c("Water", "Electricity", "Sanitation")
      )
  ) %>% 
  drop_na(population_percent) %>%
  group_by(infrastructure, u_r, iso2c) %>% 
  arrange(iso2c, infrastructure, u_r, year) %>% 
  summarize(
    decade_change = 
      last(population_percent, year) - first(population_percent, year)
  ) %>% 
  ungroup() %>% 
  ggplot(aes(decade_change)) +
  geom_histogram(binwidth = 5) +
  geom_vline(
    data = 
      . %>% 
      group_by(infrastructure, u_r) %>% 
      summarize(median = median(decade_change)),
    aes(xintercept = median, color = "median")
  ) +
  geom_vline(xintercept = 0, color = "black") +
  facet_grid(rows = vars(infrastructure), cols = vars(u_r)) +
  scale_x_continuous(
    breaks = scales::breaks_width(25),
    labels = scales::percent_format(scale = 1)
  ) +
  labs(
    title = str_glue("Distribution of change to access to infrastructure from {YEAR_MIN} - {YEAR_MAX}"),
    x = "Change to Access to Infrastructure (Percent of Population)",
    y = NULL,
    color = NULL
  )
# solution-end


# task-begin
# Print results
if (exists("q4")) q4
# task-end

# Alternate Answer:
# q3 <- 
#   q2 %>% 
#   filter(
#     iso2c %in% world_regions
#   ) %>% 
#   mutate(
#     u_r = 
#       ordered(
#         u_r, 
#         levels = c("rural", "urban"), 
#         labels = c("Rural", "Urban")
#       ),
#     infrastructure = 
#       ordered(
#         infrastructure, 
#         levels = c("water", "electricity", "sanitation"),
#         labels = c("Water", "Electricity", "Sanitation")
#       )
#   ) %>%
#   ggplot(aes(year, population_percent, color = country)) +
#   geom_line() +
#   geom_point() +
#   facet_grid(rows = vars(infrastructure), cols = vars(u_r)) +
#   scale_x_continuous(
#     breaks = scales::breaks_width(4),
#     minor_breaks = NULL
#   ) +
#   scale_y_continuous(labels = scales::percent_format(scale = 1)) +
#   labs(
#     title = "Percent of population with access to infrastructure across world regions",
#     x = "Year",
#     y = "Access to Infrastructure (Percent of Population)",
#     color = "Region",
#     captions = "Source: World Bank Data"
#   )
```

```{r}
q3.2 <- 
  q2 %>% 
  filter(
    !(iso2c %in% group_codes),
    year %in% c("2007", "2017")
  ) %>%
  mutate(
    u_r = ordered(
      u_r,
      levels = c("urban", "rural"),
      labels = c("Urban", "Rural")
    ),
    infrastructure = 
      ordered(
        infrastructure, 
        levels = c("water", "electricity", "sanitation"),
        labels = c("Water", "Electricity", "Sanitation")
      )
  ) %>% 
  drop_na(population_percent) %>%
  group_by(infrastructure, u_r, iso2c) %>% 
  arrange(iso2c, infrastructure, u_r, year) %>% 
  summarize(
    decade_change = 
      last(population_percent, year) - first(population_percent, year)
  ) %>% 
  ungroup() %>% 
  ggplot(aes(decade_change)) +
  geom_histogram(binwidth = 5) +
  geom_vline(
    data = 
      . %>% 
      group_by(infrastructure, u_r) %>% 
      summarize(median = median(decade_change)),
    aes(xintercept = median, color = "median")
  ) +
  geom_vline(xintercept = 0, color = "black") +
  facet_grid(rows = vars(infrastructure), cols = vars(u_r)) +
  scale_x_continuous(
    breaks = scales::breaks_width(25),
    labels = scales::percent_format(scale = 1)
  ) +
  theme(
    legend.position = "bottom"
  ) +
  labs(
    title = str_glue("Distribution of change to access to infrastructure from {YEAR_MIN} - {YEAR_MAX}"),
    x = "Change to Access to Infrastructure (Percent of Population)",
    y = NULL,
    color = NULL
  )
```
<!-- solution-begin -->

* Mostly positive change in the decade
* Rural change is greater
<!-- solution-end -->
